Solution of an Acoustic Transmission Inverse Problem by Extended Inversion: Theory

### 2024-02-15 repro/papers/awigrad/paper.tex
cleaned up

\section{Introduction}

@@

Then equation \ref{eqn:penderiv} can be re-written as
\[
  \langle D(T_{\sigma}[m]u)\delta m, T_{\sigma}[m]u\rangle = \langle DF[m]\delta m, 
\]
\begin{equation}
  \label{eqn:penderiv1}
   K[v_{0,\sigma}[m,u]]^T(S[m]u_{0,\sigma}[m]-d)
+
K[u_{0,\sigma}[m]]^T S[m]v_{0,\sigma}[m,u]\rangle
\end{equation}
Combine equations \ref{eqn:resderiv} and \ref{eqn:penderiv1} with
equation \ref{eqn:basederiv} to obtain
\[
  \nabla \tilde{J}_{\alpha,\sigma}[m] =
  DF[m]^T\{ K[u_{\alpha,\sigma}[m]]^T (S[m]u_{\alpha,\sigma}[m]-d)
\]
\begin{equation}
  \label{eqn:deriv}
  + \alpha^2 (
  K[v_{0,\sigma}[m, u_{\alpha,\sigma}[m]]]^T(S[m]u_{0,\sigma}[m]-d)
+ K[u_{0,\sigma}[m]]^T S[m]v_{0,\sigma}[m,u_{\alpha,\sigma}[m]]) \}.
\end{equation}

@@

\subsection{Multiple Traces}
The general case follows immediately from the single trace case, since
every operator ($S[m]$, $K[...]$, $W$, $R$) ,appearing in the right-hand side of expression
\ref{eqn:deriv} is block-diagonal (trace-by-trace), with the exception
of $DF[m]^T$. This latter is a reduction operation, i.e. a block-row
operation, so its expression is also the same. That is, equation
\ref{eqn:deriv} also correctly expresses the gradient in the
multi-trace case.

@@

The derivative of the expression on the right-hand side of
\ref{eqn:wg} is the sum of two terms, the first of which is given by
\ref{eqn:penderiv1} with $u = u_{0,\sigma}[m]$. The second is
\[
  \langle T_{\sigma}[m]u_{0,\sigma}[m], T_{\sigma}[m]D_m
  u_{0,\sigma}[m]\delta m \rangle
\]
which from identity \ref{eqn:blockderiv} is
\[
  = -\langle  T_{\sigma}[m]^TT_{\sigma}[m]u_{0,\sigma}[m], (S[m]^TS[m] + \sigma^2I)^{-1}
\]
\[
  \times ((DS[m]^{T}\delta m) (S[m]u_{0,\sigma}[m]-d) +
  S[m]^T(DS[m]\delta) m u_{0,\sigma}[m])
\]
Since
$T_{\sigma}[m]^TT_{\sigma}[m] = W(Ru_{0,\sigma}[m],-2,2)$, this is
\[
  = -\langle (DS[m]\delta m) (S[m]^TS[m] +
  \sigma^2I)^{-1}W(Ru_{0,\sigma}[m],-2,2)u_{0,\sigma}[m],
  S[m]u_{0,\sigma}[m] - d \rangle
\]
\begin{equation}
  \label{eqn:term2}
  - \langle S[m] (S[m]^TS[m] +
  \sigma^2I)^{-1}W(Ru_{0,\sigma}[m],-2,2)u_{0,\sigma}[m], (DS[m]\delta
  m)u_{0,\sigma}[m] \rangle
\end{equation}
Introduce the abbreviation
\begin{equation}
  \label{eqn:defv1}
  v_{1,\sigma}[m] = (S[m]^TS[m] + \sigma^2I)^{-1}W(Ru_{0,\sigma}[m],-2,2)u_{0,\sigma}[m].
\end{equation}
Then the RHS of equation \ref{eqn:term2bis} may be re-written as
\begin{equation}
  \label{eqn:term2bis}
  = -\langle DF[m]\delta m, K[v_{1,\sigma}[m]]^T (S[m]u_{0,\sigma}[m]-d) + 
  K[u_{0,\sigma}[m]]^T S[m] v_{1,\sigma}[m] \rangle.
\end{equation}
Recall that we have just computed the second of two terms, the sum of
which is the derivative of $\tilde{J}_{\sigma}[m]$. The first is
\ref{eqn:penderiv1} with $u = u_{0,\sigma}[m]$. Combine the two
terms to obtain
\[
  D\tilde{J}_{\sigma}[m] \delta m = \langle DF[m]\delta m,
\]
\[
  K[v_{0,\sigma}[m,u_{0,\sigma}[m]]]^T(S[m]u_{0,\sigma}[m]-d)
+
K[u_{0,\sigma}[m]]^T S[m]v_{0,\sigma}[m,u_{0,\sigma}[m]]\rangle
\]
\[
  -\langle DF[m]\delta m, K[v_{1,\sigma}[m]]^T (S[m]u_{0,\sigma}[m]-d) + 
  K[u_{0,\sigma}[m]]^T S[m] v_{1,\sigma}[m] \rangle
\]
\[
  = \langle DF[m]\delta
  m,(K[v_{0,\sigma}[m,u_{0,\sigma}[m]]]-K[v_{1,\sigma}[m]])^T(S[m]u_{0,\sigma}[m]-d)
\]
\[
  + K[u_{0,\sigma}[m]]^T S[m] (v_{0,\sigma}[m,u_{0,\sigma}[m]] -
  v_{1,\sigma}[m]) \rangle.
\]
To simplify this expression further, and reduce the computational
complexity of the derivative expression, note that the
operator-valued function $K$ is linear in its argument, according to
its definition \ref{eqn:convdef}. So the above is
\begin{equation}
  \label{eqn:wgderiv}
  = \langle DF[m]\delta m, (K[v_{\sigma}[m]]^T(S[m]u_{0,\sigma}[m]-d) 
  + K[u_{0,\sigma}[m]]^T S[m] v_{\sigma}[m] )\rangle.
\end{equation}
in which
\[
  v_{\sigma}[m] = v_{0,\sigma}[m,u_{0,\sigma}[m]] -  v_{1,\sigma}[m].
\]
The definitions \ref{eqn:defv0} and \ref{eqn:defv1} imply that
$v_{\sigma}[m]$ is the solution of the linear system
\begin{equation}
  \label{eqn:defvsigma}
  (S[m]^TS[m]+ \sigma^2I) v_{\sigma}[m] =
  (W(R(W(Ru_{0,\sigma}[m],-2,1)u_{0,\sigma}),2,0) - W(Ru_{0,\sigma}[m],-2,2))
  u_{0,\sigma}[m].
\end{equation}
That is, computation of $v_{\sigma}[m]$, hence of $D
\tilde{J}_{\sigma}[m]\delta m$, requires the solution of only one additional
linear system, beyond that necessary to compute the value of $\tilde{J}_{\sigma}[m]$.

@@

\append{Unpreconditioned AWI: Matched Source Waveform Inversion}
Matched Source Waveform Inversion (MSWI) is a close relative to
AWI, differing in the use of a simpler penalty operator $T$, defined
by
\begin{equation}
  \label{eqn:upf}
  Tu (\bx_r,t;\bx_s) = t u(\bx_r,t;\bx_s).
\end{equation}
Similarly to AWI, $m$ is estimated by minimizing
\begin{equation}
  \label{eqn:junp}
 J^{\rm unp}_{\alpha,\sigma}[u,m,d] = \frac{1}{2}(\|S[m]u - d\|_D^2 + \alpha^2\|Tu\|_U^2 + \sigma^2 \|u\|^2_U).
\end{equation}
or equivalently
\begin{equation}
  \label{eqn:jtildeunp}
 \tilde{J}^{\rm unp}_{\alpha,\sigma}[m,d] = \frac{1}{2}(\|S[m]u_{\alpha,\sigma}[m,d] - d\|_D^2 + \alpha^2\|Tu_{\alpha,\sigma}[m,d]\|_U^2 + \sigma^2 \|u_{\alpha,\sigma}\|^2_U).
\end{equation}
in which $u_{\alpha,\sigma}$ is the solution of the normal equation
\begin{equation}
  \label{eqn:normalunp}
  (S[m]^TS[m] + \alpha^2T^TT + \sigma^2I)u_{\alpha,\sigma}[m] = S[m]^Td
\end{equation}
The only difference between AWI and MSWI lies in the defintiion of the
penalty operator.

@@

Since the MSWI penalty operator $T$ is $m$-independent, it does not
appear explicilly in the the gradient
of the reduced MSWI objective $\tilde{J}^{\rm unp}_{\alpha,\sigma}$, as
given by the VPM gradient formula:
\[
  \nabla \tilde{J}^{\rm unp}_{\alpha,\sigma}[m] =
  D(S[m]u)^T_{u=u_{\alpha,\sigma}[m]}(S[m]u_{\alpha,\sigma}[m]-d)
\]
\begin{equation}
  \label{eqn:tildejunpgrad}
  = DF[m]^T(K[u_{\alpha,\sigma}[m]]^T(S[m]u_{\alpha,\sigma}[m]-d).
\end{equation}

@@

>>>>

\section{Introduction}
\section{A Penalty Version of AWI}
The adaptive kernel $u_0^{\epsilon}$ appearing in the definition \ref{eqn:tJz} is the solution of a regularized least squares problem. This observation suggests augmenting the regularized least squares objective by a penalty proportional to the right-hand side of \ref{eqn:tJz}:
\begin{equation}
  \label{eqn:Ja}
  \Ja[m,u;d] = \frac{1}{2}\left(\|S[m]u-d\|_d^2 + \epsilon^2\|u\| + \alpha^2\|tu\|_m\right).
\end{equation}
Here $\|u\|^2 = \langle u, u \rangle$ is the (unweighted) $L^2$ norm, and $\|\cdot\|_m$ is an $(m,d)$-dependent (weighted) norm: with
\[
  u_0^{\epsilon}[m,d] = \mbox{arg min}_u \Jz[m,u;d],
\]
define
\[
  Wu (\bx_r,t;\bx_s) = \sum_{\bx_r,\bx_s}\frac{1}
  {\int\,ds\,|u_0^{\epsilon}[m,d](\bx_r,s;\bx_s)|^2}u(\bx_r,t;\bx_s)
\]
and
\begin{equation}
  \label{eqn:defmnorm}
  \|u\|^2_m = \langle u, W u \rangle.
\end{equation}
With this notation,
\[
  \tJz[m;d] = \frac{1}{2}\|tu^{\epsilon}_0[m;d]\|^2_m
\]
In the definition of $\tJz$, a quadratic is evaluated at the minimizer of $\Jz[m,u;d]$. This suggests the possibility of evaluating the quadratic objective $\Ja[m,u;d]$ at its own minimizer:
\begin{equation}
  \label{eqn:tJa}
  \tJa[m;d] = \min_u \Ja[m,u,d] = \Ja[m,\ua[m;d];d]
\end{equation}
This is the so-called variable projection reduction of the quadratic $\Ja$.
Note that the right-hand side of the definition of $\tJz$ is the last term on the right-hand side of $\Ja$, scaled by $\alpha^{-2}$ and evaluated at the minimizer of the sum of the first two terms, that is, $\Ja$ for $\alpha=0$.

\begin{theorem}
  \label{thm:basic}
Suppose that $S \in {\cal L}(U,D)$, the is a bounded linear maps from (Hilbert space) domain $U$ (``model space'') with inner product $\langle \cdot, \cdot \rangle_m$ to range $D$ with inner product $\langle \cdot, \cdot \rangle_d$. Assume that $S$ is coercive, that is, there exists $\epsilon >0$ so that $\|Su\|_d \ge \epsilon\|u\|_m$ for $u \in U$.
Suppose also that $A \in {\cal L}(U,U)$, and that $d \in D$. Define the function $J_{\alpha}$ on $U$ by 
\begin{equation}
  \label{eqn:eq1}
  J_{\alpha}(u) = \frac{1}{2}(\|Su-d\|^2_d + \alpha^2\|Au\|_m^2).
\end{equation}
Then for $\alpha \ge 0$, $J_{\alpha}$ possesses a unique minimizer $u_{\alpha} \in U$, for which
\begin{equation}
  \label{eqn:eq0}
  \tilde{J}_{\alpha} = \min_{u \in U} J_{\alpha}(u) = J_{\alpha}(u_{\alpha}),
\end{equation}
and
\begin{equation}
  \label{eqn:eq4}
  \lim_{\alpha \rightarrow 0} \frac{1}{\alpha^2}  (\tilde{J}_{\alpha}-\tilde{J}_0).= \frac{1}{2}\|Au_0\|^2
\end{equation}
This limit is uniform over bounded sets of $(S, A) \in {\cal L}(U,D) \times {\cal L}(U, U)$.
\end{theorem}

For completeness, the proof is given in an appendix.

This relation will be applied in the context of a model space $U$ possessing a reference Hilbert norm $\|\cdot\|$ and a family of equivalent norms $\|\cdot\|_m$ depending on a parameter $m \in M$, in which $M$ is an open subset of a Banach space $X$. The dependence is through a an operator-valued function $W$ on $M$, taking values in cone ${\cal B}(U)$ of bounded self-adjoint positive-definite operators on $U$. On the other hand, we assume that the norm on $D$ is fixed, independent of $M$, and treat it as a reference norm.

\begin{theorem}
  \label{thm:uniform}
Assume that $W: M \rightarrow {\cal B}(U)$ is of class $C^1$ and uniformly positive definite, and that the norm $\|\cdot\|_m$ is given by
\begin{equation}
  \label{eqn:normdef}
  \|u\|_m^2 = \langle u, W[m] u \rangle.
\end{equation}
Suppose also that $S: M \rightarrow {\cal L}(U,D)$ is of class $C^1$ and uniformly coercive, and that $A \in {\cal L}(U,U)$ as before. For $\alpha \ge 0$, define $\tJa: M \rightarrow \bR$ by
\[
  \tilde{J}_{\alpha}[m] = \min_{u \in U}\frac{1}{2}(\|S[m]u-d\|^2_d + \alpha^2\|Au\|_m^2)
\]
\begin{equation}
  \label{eqn:tjadef}
  = \frac{1}{2}(\|S[m]u_{\alpha}[m]-d\|^2_d + \alpha^2\|Au_{\alpha}[m]\|_m^2)
\end{equation}
where
\begin{equation}
  \label{eqn:ua}
  u_{\alpha}[m] = \mbox{arg min}_{u \in U}\frac{1}{2}(\|S[m]u-d\|^2_d + \alpha^2\|Au\|_m^2)
\end{equation}
for $m \in M$. Then $\tilde{J}_{\alpha}$ is of class $C^1(M)$, and
\begin{equation}
  \label{eqn:lim}
  \lim_{\alpha \rightarrow 0} \frac{1}{\alpha^2}  (\tJa[m]-\tJz[m]).= \frac{1}{2}\|A\uz[m]\|_m^2.
\end{equation}
The limit is in the sense of $C^1(M)$, that is, both the left-hand side and its $m$ derivative converge to the right-hand side and its $m$ derivative, respectively. Moreover $u_{\alpha}$ is of class $C^1(M,U)$.
\end{theorem}

\begin{proof} All conclusions but the last follow immediately from Theorem \ref{thm:basic} and the uniformity assumptions on $S$ and $W$.

Since the norm in $D$ is fixed for the discussion, drop the subscript and treat it as a reference norm. Denote the adjoint with respect to the norms $\|\cdot\|_m$ on $U$, $\|\cdot\|_d = \|\cdot\|$ on $D$ by a superscript $*$, and with respect to $\|\cdot\|$ on $U$ by a superscript $T$. The normal equation defining $\ua$ is
  \[
    (S[m]^*S[m] + \alpha^2 A[m]^*A) \ua[m] = S[m]^*d
  \]
  \begin{equation}
    \label{eqn:normal}
= (W[m]^{-1}S[m]^TS[m] + \alpha^2 W[m]^{-1}A^TW[m]A)\ua[m]= W[m]^{-1}S[m]^Td
\end{equation}
or equivalently
\begin{equation}
  \label{eqn:normal1}
  (S[m]^TS[m] + \alpha^2 A^TW[m]A)\ua[m] = S[m]^Td
\end{equation}
Since $S$ is uniformly coercive, and both $S$ and $W$ are of class $C^1$, the normal operator on the left hand side is boundedly invertible and of class $C^1$. Therefore the solution $\ua$ is of class $C^1$.

[need to add proof of $C^1$ convergence.]
\end{proof}

Apply this relation in the context of the AWI penalty problem defined earlier. $M$ is an open set in $C^k(\Omega, \bR^2)$. $U$ is $\oplus_{\bx_r,\bx_s}L^2([-T_u,T_u])$, $D$ is the data space $\oplus_{\bx_r,\bx_s}L^2([0,T])$,. $S$ is the regularized modeling operator $(S[m],\epsilon I)^T$, $d$ is replaced by the $(d,0) \in D \times U$,  $W$ is the weight operator on $U$ defined in \ref{eqn:defmnorm}.

As noted before [but it wasn't - fix!], stationary points of $\tJz$
are $\lambda(w_*)$-close to stationary points of the travel time
mean-square error, in the single-arrival case. The result of this
section implies that stationary points of $\tJa$ are close to those of
$\tJz$. Therefore, at least for small $\alpha$, minimization of $\tJa$
produces acoustic models that closely match the traveltimes predicted
in single-arrival data.

@@

Then equation \ref{eqn:penderiv} can be re-written as
\[
  \langle D(T_{\sigma}[m]u)\delta m, T_{\sigma}[m]u\rangle = \langle DF[m]\delta m, 
\]
\begin{equation}
  \label{eqn:penderiv1}
   K[v_{0,\sigma}[m,u]]^T(S[m]u_{0,\sigma}[m]-d)
+
K[u_{0,\sigma}[m]]^T S[m]v_{0,\sigma}[m,u]\rangle
\end{equation}
Combine equations \ref{eqn:resderiv} and \ref{eqn:penderiv1} with
equation \ref{eqn:basederiv} to obtain
\[
  \nabla \tilde{J}_{\alpha,\sigma}[m] =
  DF[m]^T\{ K[u_{\alpha,\sigma}[m]]^T (S[m]u_{\alpha,\sigma}[m]-d)
\]
\begin{equation}
  \label{eqn:tildejgrad}
  + \alpha^2 (
  K[v_{0,\sigma}[m, u_{\alpha,\sigma}[m]]]^T(S[m]u_{0,\sigma}[m]-d)
+ K[u_{0,\sigma}[m]]^T S[m]v_{0,\sigma}[m,u_{\alpha,\sigma}[m]]) \}.
\end{equation}

@@

\subsection{Multiple Traces}
The general case follows immediately from the single trace case, since
every operator ($S[m]$, $K[...]$, $W$, $R$) ,appearing in the right-hand side of expression
\ref{eqn:tildejgrad} is block-diagonal (trace-by-trace), with the exception
of $DF[m]^T$. This latter is a reduction operation, i.e. a block-row
operation, so its expression is also the same. That is, equation
\ref{eqn:tildejgrad} also correctly expresses the gradient in the
multi-trace case.

@@

The derivative of the expression on the right-hand side of
\ref{eqn:wg} is the sum of two terms, the first of which is given by
\ref{eqn:penderiv1} with $u = u_{0,\sigma}[m]$. The second is
\[
  \langle T_{\sigma}[m]u_{0,\sigma}[m], T_{\sigma}[m]D_m
  u_{0,\sigma}[m]\delta m \rangle
\]
which from identity \ref{eqn:blockderiv} is
\[
  = -\langle  T_{\sigma}[m]^TT_{\sigma}[m]u_{0,\sigma}[m], (S[m]^TS[m] + \sigma^2I)^{-1}
\]
\[
  \times ((DS[m]^{T}\delta m) (S[m]u_{0,\sigma}[m]-d) +
  S[m]^T(DS[m]\delta) m u_{0,\sigma}[m])
\]
Since
$T_{\sigma}[m]^TT_{\sigma}[m] = W(Ru_{0,\sigma}[m],-2,2)$, this is
\[
  = -\langle (DS[m]\delta m) (S[m]^TS[m] +
  \sigma^2I)^{-1}W(Ru_{0,\sigma}[m],-2,2)u_{0,\sigma}[m],
  S[m]u_{0,\sigma}[m] - d \rangle
\]
\begin{equation}
  \label{eqn:term2}
  - \langle S[m] (S[m]^TS[m] +
  \sigma^2I)^{-1}W(Ru_{0,\sigma}[m],-2,2)u_{0,\sigma}[m], (DS[m]\delta
  m)u_{0,\sigma}[m] \rangle
\end{equation}
Introduce the abbreviation
\begin{equation}
  \label{eqn:defv1}
  v_{1,\sigma}[m] = (S[m]^TS[m] + \sigma^2I)^{-1}W(Ru_{0,\sigma}[m],-2,2)u_{0,\sigma}[m].
\end{equation}
Then the RHS of equation \ref{eqn:term2bis} may be re-written as
\begin{equation}
  \label{eqn:term2bis}
  = -\langle DF[m]\delta m, K[v_{1,\sigma}[m]]^T (S[m]u_{0,\sigma}[m]-d) + 
  K[u_{0,\sigma}[m]]^T S[m] v_{1,\sigma}[m] \rangle.
\end{equation}
Recall that we have just computed the second of two terms, the sum of
which is the derivative of $\tilde{J}_{\sigma}[m]$. The first is
\ref{eqn:penderiv1} with $u = u_{0,\sigma}[m]$. Combine the two
terms to obtain
\[
  D\tilde{J}_{\sigma}[m] \delta m = \langle DF[m]\delta m,
\]
\[
  K[v_{0,\sigma}[m,u_{0,\sigma}[m]]]^T(S[m]u_{0,\sigma}[m]-d)
+
K[u_{0,\sigma}[m]]^T S[m]v_{0,\sigma}[m,u_{0,\sigma}[m]]\rangle
\]
\[
  -\langle DF[m]\delta m, K[v_{1,\sigma}[m]]^T (S[m]u_{0,\sigma}[m]-d) + 
  K[u_{0,\sigma}[m]]^T S[m] v_{1,\sigma}[m] \rangle
\]
\[
  = \langle DF[m]\delta
  m,(K[v_{0,\sigma}[m,u_{0,\sigma}[m]]]-K[v_{1,\sigma}[m]])^T(S[m]u_{0,\sigma}[m]-d)
\]
\[
  + K[u_{0,\sigma}[m]]^T S[m] (v_{0,\sigma}[m,u_{0,\sigma}[m]] -
  v_{1,\sigma}[m]) \rangle.
\]
To simplify this expression further, and reduce the computational
complexity of the derivative expression, note that the
operator-valued function $K$ is linear in its argument, according to
its definition \ref{eqn:convdef}. So the above is
\begin{equation}
  \label{eqn:wgderiv}
  = \langle DF[m]\delta m, (K[v_{\sigma}[m]]^T(S[m]u_{0,\sigma}[m]-d) 
  + K[u_{0,\sigma}[m]]^T S[m] v_{\sigma}[m] )\rangle.
\end{equation}
in which
\[
  v_{\sigma}[m] = v_{0,\sigma}[m,u_{0,\sigma}[m]] -  v_{1,\sigma}[m].
\]
The definitions \ref{eqn:defv0} and \ref{eqn:defv1} imply that
$v_{\sigma}[m]$ is the solution of the linear system
\[
    (S[m]^TS[m]+ \sigma^2I) v_{\sigma}[m] =
\]
\begin{equation}
  \label{eqn:defvsigma}
  (W(R(W(Ru_{0,\sigma}[m],-2,1)u_{0,\sigma}),2,0) - W(Ru_{0,\sigma}[m],-2,2))
  u_{0,\sigma}[m].
\end{equation}
That is, computation of $v_{\sigma}[m]$, hence of $D
\tilde{J}_{\sigma}[m]\delta m$, requires the solution of only one additional
linear system, beyond that necessary to compute the value of $\tilde{J}_{\sigma}[m]$.

@@

\append{Unpreconditioned AWI: Matched Source Waveform Inversion}
Matched Source Waveform Inversion (MSWI) is a close relative to
AWI, differing in the use of a simpler penalty operator $T$, defined
by
\begin{equation}
  \label{eqn:upf}
  Tu (\bx_r,t;\bx_s) = t u(\bx_r,t;\bx_s).
\end{equation}
Similarly to AWI, $m$ is estimated by minimizing
\begin{equation}
  \label{eqn:junp}
 J^{\rm unp}_{\alpha,\sigma}[u,m,d] = \frac{1}{2}(\|S[m]u - d\|_D^2 + \alpha^2\|Tu\|_U^2 + \sigma^2 \|u\|^2_U).
\end{equation}
or equivalently
\begin{equation}
  \label{eqn:jtildeunp}
 \tilde{J}^{\rm unp}_{\alpha,\sigma}[m,d] = \frac{1}{2}(\|S[m]u_{\alpha,\sigma}[m,d] - d\|_D^2 + \alpha^2\|Tu_{\alpha,\sigma}[m,d]\|_U^2 + \sigma^2 \|u_{\alpha,\sigma}\|^2_U).
\end{equation}
in which $u_{\alpha,\sigma}$ is the solution of the normal equation
\begin{equation}
  \label{eqn:normalunp}
  (S[m]^TS[m] + \alpha^2T^TT + \sigma^2I)u_{\alpha,\sigma}[m] = S[m]^Td
\end{equation}
The only difference between AWI and MSWI lies in the absence of
scaling by the reciprocal of $Ru_{0,\sigma}[m]$ in the definition of the MSWI
penalty operator.

@@

Since the MSWI penalty operator $T$ is $m$-independent, it does not
appear explicilly in the the gradient
of the reduced MSWI objective $\tilde{J}^{\rm unp}_{\alpha,\sigma}$, as
given by the VPM gradient formula:
\[
  \nabla \tilde{J}^{\rm unp}_{\alpha,\sigma}[m] =
  D_m(S[m]u)^T_{u=u_{\alpha,\sigma}[m]}(S[m]u_{\alpha,\sigma}[m]-d)
\]
\begin{equation}
  \label{eqn:tildejunpgrad}
  = DF[m]^TK[u_{\alpha,\sigma}[m]]^T(S[m]u_{\alpha,\sigma}[m]-d).
\end{equation}

Note that the right-hand side of equation \ref{eqn:tildejunpgrad} is
exactly the first term in the right-hand side of equation
\ref{eqn:tildejgrad}. That is, the gradient of the AWI objective
function is the gradient of the MSWI objective function, modified by
terms that account for the $m$-dependence of the AWI penalty operator $T_{\sigma}[m]$.

@@

\append{$\alpha \rightarrow 0$ scaled limit of penalty function}
Define
\begin{equation}
  \label{eqn:eq0}
  \tJa = \min_u J_{\alpha}(u) = J_{\alpha}(\ua)
\end{equation}
Since $S$ is positive definite, the minimum is well-defined for any $\alpha \ge 0$, and the minimizer $\ua \in U$ satisfies the normal equation.

The claim to be established is that
\begin{equation}
  \label{eqn:eq4}
  \lim_{\alpha \rightarrow 0} \frac{1}{\alpha^2}  (\tJa-\tJz).= \frac{1}{2}\|A\uz\|_m^2
\end{equation}

To see this, use the normal equation to write
\[
  \ua = (S^TS + \alpha^2 A^TA)^{-1}S^Td = (S^TS + \alpha^2 A^TA)^{-1}S^TS \uz
\]
\begin{equation}
  \label{eqn:eq2}
  = \uz - \alpha^2 (S^TS + \alpha^2 A^TA)^{-1}A^TA\uz = \uz-\alpha^2 \va
\end{equation}
Note that $\va = (S^TS + \alpha^2 A^TA)^{-1}A^TA\uz$ is uniformly bounded in $\alpha \ge 0$.

Stuff the RHS of equation \ref{eqn:eq2} into the definition \ref{eqn:eq1} to obtain
\begin{equation}
  \label{eqn:eq3}
  \tJa = \frac{1}{2}(\|S\uz-d\|_d^2 - 2 \alpha^2\langle S\uz-d, S\va\rangle_d + \alpha^2\|A\uz\|_m^2 + O(\alpha^4))
\end{equation}
The second term on the RHS of equation \ref{eqn:eq3} vanishes thanks to the normal equation, and the first term is precisely $\tJz$. The conclusion \ref{eqn:eq4} follows.

### 2024-02-15 repro/papers/discreptheory/paper.tex
cleaned up

\section{Variable Projection}
The main theoretical device used in the proofs of our main results on 
extended inversion is Variable Projection reduction of the penalty objective $\Ja$
(equation \ref{eqn:penerr}) to a function $\tJa$ of 
$m$ alone, by minimization over $w$:
\begin{equation}
  \label{eqn:redexp}
  \tJa[m;d] = \inf_w \Ja[m,w;d]
\end{equation}

@@

This is as far as one can go without more concrete assumptions on $F$
and $A$. If the reduced
objective $\tJa$ possesses a large (in some sense) domain of convexity for a range of $\alpha$, then the discrepancy algorithm sketched
here can efficiently drive $\alpha$ to near its largest feasible
value, thus giving the best resolution near the global minimizer. That
program has been carried out for the simple cartoon problem of this
paper, and partly for some other more protypical inverse wave problems \cite[]{FuSymes2017discrep}.

>>>>

\section{Variable Projection}
The main theoretical device used in the proofs of our main results on 
extended inversion is Variable Projection reduction of the penalty objective $\Ja$
(equation \ref{eqn:jdef}) to a function $\tJa$ of 
$m$ alone, by minimization over $w$:
\begin{equation}
  \label{eqn:redexp}
  \tJa[m;d] = \inf_w \Ja[m,w;d]
\end{equation}

@@

This is as far as one can go without more concrete assumptions on $F$
and $A$. If the reduced
objective $\tJa$ possesses a large (in some sense) domain of convexity for a range of $\alpha$, then the discrepancy algorithm sketched
here can efficiently drive $\alpha$ to near its largest feasible
value, thus giving the best resolution near the global minimizer. That
program has been carried out for the simple cartoon problem of this
paper, and partly for some other more protypical inverse wave problems \cite[]{FuSymes2017discrepancy}.
